---
title: "Assignment_2_MSB1015"
author: "Wiep Stikvoort"
date: "October 12, 2019"
output: html_notebook
---
rrdkit? <- r package
rcdk vignet, multilinear regression model for boiling points
github repository set up for assignment 2

get the smiles and boiling points with the wikidataqueryserviceR package
parse them with the rrdkit or the rcdk

First, we need to load the packages needed for the query and the computations for the regression model. 
```{r setup, installing packages, echo=FALSE}
install.packages('SPARQL')
library(SPARQL)
install.packages('rJava')
library(rJava)
install.packages('rcdk')
library(rcdk)
install.packages('pls')
library(pls)

library(RCurl)

install.packages('WikidataQueryServiceR')
library(WikidataQueryServiceR)
install.packages('caTools')
library('caTools')
library('ggplot2')

```


This is the correct query:
``` {r}
query <- query_wikidata('SELECT DISTINCT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?CC
WHERE {   
?comp wdt:P31/wdt:P279* wd:Q41581 ;       
wdt:P233 ?CC  ;
p:P2102 [ ps:P2102 ?bp ;  
psv:P2102/wikibase:quantityUnit  ?bpUnit         ] .  
SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". } }')
```

There are still some compounds with their boiling points in degree celcius or fahrenheit. These need to be changed to kelvin:

```{r}
fahrenheit_kelvin <- function(fahrenheit) {
  kelvin <- (fahrenheit + 459.67) * (5/9)
  return(kelvin)
}
celsius_kelvin <- function(celcius) {
  kelvin <- celcius + 293.15
  return(kelvin)
}

for (i in 1:length(query$bpUnitLabel)) {
  if(query$bpUnitLabel[i] == 'degree Celsius'){
      query$bp[i] <- celsius_kelvin(query$bp[i])
      query$bpUnitLabel[i] <- 'kelvin'
      }
  else if (query$bpUnitLabel[i] == 'degree Fahrenheit'){
     query$bp[i] <- fahrenheit_kelvin(query$bp[i])
     query$bpUnitLabel[i] <- 'kelvin'
  }
}

plot.default(query$bp)

```

Splitting the data into testset and trainingset
```{r}
sample = sample.split(query, SplitRatio = 3/4)
training_set = subset(query, sample == TRUE)
test_set = subset(query, sample == FALSE)
```

new method for splitting into test and training set
```{r}
data <- sort(sample(nrow(query), nrow(query)*.8))
training_set<-query[data,]

#creating test data set by not selecting the output row values
test_set<-query[-data,]
```


descriptors 
```{r}
parsed_smiles <- parse.smiles(training_set$CC, kekulise=TRUE) # changes it into a format IAtomContainer

descNames <- c(
'org.openscience.cdk.qsar.descriptors.molecular.FragmentComplexityDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.MDEDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.AtomCountDescriptor')

descs <- eval.desc(parsed_smiles, descNames)
training <- cbind(training_set, descs)

```

``` {r}

model <- plsr(bp ~ WPATH + fragC + apol + WPATH + WPOL + MDEC.12 + nAtom, data = training, validation = "LOO") 

plot(model, asp = 1, line = TRUE)


parsed_smiles_test <- parse.smiles(test_set$CC, kekulise = TRUE)
descs_test <- eval.desc(parsed_smiles_test, descNames)

plot(RMSEP(model), legendpos = "topright")

test <- cbind(test_set, descs_test)
error <- RMSEP(model, newdata = test)
plot(error, legendpos = "topright")

predict <- predict(model, ncomp = 3, newdata = test) # 
plot(predict, test_set$bp)
corcoef <- cor(predict, test_set$bp)
print(error)
```

``` {r}
query_amino_acids <- query_wikidata('SELECT DISTINCT ?comp ?compLabel ?mass ?massUnit ?massUnitLabel ?CC
WHERE {   
?comp wdt:P31/wdt:P279* wd:Q8066 ;       
wdt:P233 ?CC  ;
p:P2067 [ ps:P2067 ?mass ;  
psv:P2067/wikibase:quantityUnit  ?massUnit         ] .  
SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". } }')
```
All compound masses were expressed in the atomic mass unit, so no need for conversions. 

```{r}
data_aa <- sort(sample(nrow(query_amino_acids), nrow(query_amino_acids)*.8))
training_set_aa<-query_amino_acids[data_aa,]

#creating test data set by not selecting the output row values
test_set_aa<-query_amino_acids[-data_aa,]
```

Descriptors 
```{r}
parsed_smiles_aa <- parse.smiles(training_set_aa$CC, kekulise=TRUE) # changes it into a format IAtomContainer

descNames <- c(
'org.openscience.cdk.qsar.descriptors.molecular.FragmentComplexityDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.MDEDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.AtomCountDescriptor')

descs_aa <- eval.desc(parsed_smiles_aa, descNames)
training_aa <- cbind(training_set_aa, descs_aa)

```

``` {r}

model_aa <- plsr(mass ~ WPATH + fragC + apol + WPATH + WPOL + MDEC.12 + nAtom, data = training_aa, validation = "LOO") 

plot(model_aa, asp = 1, line = TRUE)


parsed_smiles_test_aa <- parse.smiles(test_set_aa$CC, kekulise = TRUE)
descs_test_aa <- eval.desc(parsed_smiles_test_aa, descNames)

plot(RMSEP(model_aa), legendpos = "topright")

test_aa <- cbind(test_set_aa, descs_test_aa)
error_aa <- RMSEP(model_aa, newdata = test_aa)
plot(error_aa, legendpos = "topright")

predict_aa <- predict(model_aa, ncomp = 3, newdata = test_aa) # 
plot(predict_aa, test_set_aa$bp)
corcoef_aa <- cor(predict_aa, test_set_aa$mass)
print(error_aa)
```